\documentclass{memoir}

\title{How to write a Report\\ for the Project of Distributed Systems}

\author{Prof. Marino Miculan\\ DMIF, University of Udine, Italy}

\date{Version 0.2, \today}

\begin{document}


%\begin{titlingpage}
\maketitle
\begin{abstract}
The aim of this document is to describe how to write a report for the project-exam for the course Distributed Systems, at the University of Udine.
It is also a guideline for the development of the project itself.

This document is very brief and succint, and it is by no means comprehensive of all the informations that should be given about a software project. You are invited to supplement these guidelines as needed to best describe your work.
\end{abstract}
%\end{titlingpage}

\chapter{Introduction}\label{ch:intro}

In this chapter you describe the main problem, and an idea of the solution.
It is not necessary to be very detailed or formal, but it is important to explain which are the main aims and issues from the point of view of Distributed Systems:
\begin{itemize}
\item A description of the application.
\item The overall structure of the implementation: how resources are deployed, which are the players, the r\^oles.
\item The distributed system features (and the transparencies) and algorithms you intend to implement.
\item Your plan for testing the system.
\item A schedule for how you plan to carry our your design and implementation
\end{itemize}



\chapter{Analysis}\label{ch:analysis}

In this chapter, we describe in detail functional and non-functional requirements of a solution for the problem.

\section{Functional requirements}
Which functions must be offered to users / other programs?  Which are the input data and the output data? Which is the expected effect? 

\section{Non functional requirements}
Everything about mode and transparencies: availability, mobility, security, fault tolerance, etc.

Are there execution time bounds? Minimum data rates?

If requested, specific platforms/languages/middlewares requirements for the implementation can be decided here. (E.g.: if the project is on a SOA, we may request that functions are offered via SOAP or RESTful services). 



\chapter{Project}

This chapter is devoted to the description of the general architectures, and specific algorithms.

\section{Logical architecture}
Describe the components of your systems: modules/objects/components/services.
For each component, describe the functionalities it implements, and by who is used.

\section{Protocols and algorithms}
Communication between components.  UML sequence diagrams go here.

Also, put here a detailed description of distributed algorithms used to solve specific problems of the project.

\section{Physical architecture and deployment}
Which nodes and platforms involved, and where each component is deployed.

\section{Development plan}
Since it is difficult to predict just how hard implementing a new system will be, you should formulate as a set of ``tiers,'' where the basic tier is something you’re sure you can complete, and the additional tiers add more features, at both the application and the system level.

\chapter{Implementation}

Details about the implementation: every choice about platforms, languages, software/hardware, middlewares, which has not been decided in the requirements.


Important choices about implementation should be described here; e.g., peculiar data structures.


\chapter{Validation}

Check if requirements from Chapter~\ref{ch:analysis} have been fulfilled.
Quantitative tests (simulations) and screenshots of the interfaces are put here.


\chapter{Conclusions}

What has been done with respect to what has been promised in Chapters~\ref{ch:intro} and \ref{ch:analysis}, and what is left out.

\appendix

\chapter{Appendix}

In the Appendix you can put code snippets, snapshots, installation instructions, etc.


\chapter*{Evaluation}
Your system will be judged mainly on how it operates as a distributed system. The primary evaluation will be according to whether your system has the following attributes:
\begin{itemize}
\item  It should be an interesting distributed system, making use of some of the algorithms we have covered in class for distributed synchronization, replication, fault tolerance and recovery, security, etc.
\item The software should be well designed and well implemented, in terms of the overall architecture and the detailed realization.
\item You should devise and apply systematic testing procedures, at both the unit and systems levels.
\item The system should operate reliably and with good performance, even in the face of failures.
\end{itemize}
Important, but secondary considerations include:
\begin{itemize}
\item Time taken to do the project (the sooner the better, but do not miss details in order to end sooner)
\item  How nice is the application's appearance: does it have a nice interface or a compelling visual display?
\end{itemize}

\end{document}